Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-09-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
I am using a draft document class so that I can compile it without errors
I have changed $\bm{u}$ to $\mathbf{u}$ (I think this style makes the vectors stand out more. But you can change back if you disagree)

I have changed $\mathbf{uu}$ to $\mathbf{u}\otimes\mathbf{u}$ for clarity

I have removed your ending to the Model section and replaced it with one which cites the appendix

I have removed your appendix and replaced it with mine, in a different file to avoid conflicts. I will now continue to work on the appendix and hopefully won't make many more changes to the main file

\documentclass{ametsoc}

\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{siunitx}

@@

We use the finite volume model from \citet{weller-shahrokhi2014} which details a discretisation of the fully-compressible Euler equations, given by
\begin{subequations}
\begin{align}
	\text{Momentum} &\ &\  	\frac{\partial \rho \bm{u}}{\partial t} + \nabla \cdot \rho \bm{uu} &= \rho \bm{g} - c_p \rho \theta \nabla \Pi - \mu \rho \bm{u} \label{eq:momentum} \\
	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \bm{u} &= 0 \\
	\text{Thermodynamic equation} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \bm{u} \theta &= 0 \\
	\text{Ideal gas law} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
\end{align}
\end{subequations}
where \(\rho\) is the density, \(\bm{u}\) is the velocity field, \(\bm{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.  \(\mu\) is a dimensionless damping function used for the sponge layer in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.

The model is operated in a two-dimensional $x$--$z$ slice configuration without Coriolis forces.  The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centroids and the covariant component of velocity at cell faces.
A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.

For tracer advection tests, the advection equation is solved in flux form, given by
\begin{align}
\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0
\end{align}
where the tracer density, $\phi$, is interpolated onto cell faces using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
\begin{subequations}
\begin{align}
	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
\end{align}
\end{subequations}
where \(f(\phi^{(n)}) = - \nabla \cdot (\bm{u} \phi^{(n)})\) at time level \(n\).

@@

% APPENDICES
\appendix[A]
\appendixtitle{Non-orthogonal formulation}

\TODO{Hilary will redraft this appendix}
A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.  \citet{weller-shahrokhi2014} use a Hodge dual operator, \(H\), to map from co-variant to contra-variant velocity components:
\begin{align}
	U = HV
\end{align}
where \(U = \rho\bm{u}\cdot\bm{S}_f\) and \(V=\rho\bm{u}\cdot\bm{d}_f\).  \(\bm{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\bm{d}_f\) is the vector between neighbouring cell centres.

In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).  However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.  Improved stability is achieved by splitting \(H\) into the orthogonal part and the non-orthogonal correction: 
\begin{align}
	H = H_c + H_\mathrm{corr}
\end{align}
where the diagonal matrix \(H_c=|\bm{S}_f|/|\bm{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).  The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.

>>>>

\documentclass[draft]{ametsoc}

\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage[title]{appendix}

@@

We use the finite volume model from \citet{weller-shahrokhi2014} which details a discretisation of the fully-compressible Euler equations, given by
\begin{subequations}
\begin{align}
	\text{Momentum} &\ &\  	\frac{\partial \rho \mathbf{u}}{\partial t} + \nabla \cdot \rho \mathbf{u}\otimes\mathbf{u} &= \rho \mathbf{g} - c_p \rho \theta \nabla \Pi - \mu \rho \mathbf{u} \label{eq:momentum} \\
	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \mathbf{u} &= 0 \\
	\text{Thermodynamic equation} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \mathbf{u} \theta &= 0 \label{eq:theta} \\
	\text{Ideal gas law} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
\end{align}
\end{subequations}
where \(\rho\) is the density, \(\mathbf{u}\) is the velocity field, \(\mathbf{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.  \(\mu\) is a dimensionless damping function used for the sponge layer in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.

The model is operated in a two-dimensional $x$--$z$ slice configuration without Coriolis forces.  The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centroids and the covariant component of velocity at cell faces.
\input{model}
%A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.

%For tracer advection tests, the advection equation is solved in flux form, given by
%\begin{align}
%\partial \phi / \partial t + \nabla \cdot \left( \mathbf{u} \phi \right) = 0
%\end{align}
%where the tracer density, $\phi$, is interpolated onto cell faces using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
%The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
%\begin{subequations}
%\begin{align}
%	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
%	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
%	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
%\end{align}
%\end{subequations}
%where \(f(\phi^{(n)}) = - \nabla \cdot (\mathbf{u} \phi^{(n)})\) at time level \(n\).

@@

\begin{appendices}
\input{WellerAppendix}
\end{appendices}

%% APPENDICES
%\appendix[A]
%\appendixtitle{Non-orthogonal formulation}

%\TODO{Hilary will redraft this appendix}
%A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.  \citet{weller-shahrokhi2014} use a Hodge dual operator, \(H\), to map from co-variant to contra-variant velocity components:
%\begin{align}
%	U = HV
%\end{align}
%where \(U = \rho\mathbf{u}\cdot\mathbf{S}_f\) and \(V=\rho\mathbf{u}\cdot\mathbf{d}_f\).  \(\mathbf{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\mathbf{d}_f\) is the vector between neighbouring cell centres.

%In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).  However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.  Improved stability is achieved by splitting \(H\) into the orthogonal part and the non-orthogonal correction: 
%\begin{align}
%	H = H_c + H_\mathrm{corr}
%\end{align}
%where the diagonal matrix \(H_c=|\mathbf{S}_f|/|\mathbf{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).  The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.
